hormones

### 2020-02-03 analysis/07_hormones.Rmd
now with pCA and new correlations

knitr::opts_chunk$set(fig.path = '../figures/hormones/',message=F, warning=FALSE)
```

@@

  hormones %>% 
    filter(study == "characterization",
           hormone %in% c(myhormone))  %>% 
    droplevels() %>% 
  ggplot(aes(x = as.numeric(treatment), y = plasma_conc)) +
        geom_smooth(aes(colour = sex)) +
    geom_boxplot(aes(fill = treatment, alpha = sex)) +
    mytheme() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
          legend.position = "none") +
    scale_fill_manual(values = colorscharmaip) +
    scale_color_manual(values = sexcolors) +
    labs(y = myylab, x = NULL) +
    guides(fill = guide_legend(order=1),
         color = guide_legend(order=2)) +
    scale_alpha_manual(values = c(0.75,1)) +
    scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9),
                       labels = charlevels)
  }

@@

## for rechelle

```{r forrechelle, fig.height= 4, fig.width=6}

rechelle <- hormones %>%
  dplyr::filter(hormone %in% c("prolactin", "corticosterone"),
                treatment %in% c("lay", "inc.d9", "hatch", "n9")) %>%
  dplyr::select(-icons, -music, -iconpath)

rechelle$hormone <- factor(rechelle$hormone, levels = c("prolactin", "corticosterone"))

prl.rechelle <- rechelle %>% filter(hormone == "prolactin")   %>%  droplevels()
cort.rechelle <- rechelle %>% filter(hormone == "corticosterone")   %>%  droplevels()

TukeyHSD(aov(plasma_conc ~  treatment, data=prl.rechelle))
TukeyHSD(aov(plasma_conc ~  treatment, data=cort.rechelle))

rechelleplot <- function(mydf, myylab){

  p <- ggplot(mydf, aes(x = treatment, y = plasma_conc, fill = sex)) +
  geom_point(aes(color = sex), position = position_dodge(0.9), size = 2)+
  geom_boxplot(alpha = 0.5, outlier.shape = NA) +
  scale_color_manual(name = "Sex", values = c("gray50", "maroon"))+
  scale_fill_manual(name = "Sex", values = c("gray50", "maroon"))+

  theme(axis.text.x = element_text(angle = 45, hjust = 1),
         legend.position = "none",
          strip.text = element_blank()) +
    theme_classic()+
  theme(text = element_text(size= 15)) +
    labs(x = "Stage", y = myylab) 
  return(p)

}

#PRL
rechelleplot(prl.rechelle, "Prolactin (ng/mL)")
rechelleplot(prl.rechelle, "Prolactin (ng/mL)") + 
  scale_y_continuous(breaks = c(0,25,50,75,100, 125), limits = c(-20,125)) +
  geom_signif(comparisons=list(c("lay", "hatch")), annotations= "p < 0.001", y_position = 100, tip_length = 0.05, vjust = 0, color= "black" , textsize = 5 ) +
  geom_signif(comparisons=list(c("lay", "n9")), annotations= "p < 0.001", y_position = 110, tip_length = 0.05, vjust = 0, color= "black" , textsize = 5 ) +
  geom_signif(comparisons=list(c("hatch", "n9")), annotations= "p < 0.001", y_position = 120, tip_length = 0.05, vjust = 0, color= "black" , textsize = 5 ) +
  geom_signif(comparisons=list(c("inc.d9", "hatch")), annotations= "p < 0.001", y_position = -5, tip_length = -0.05, vjust = 2, color= "black" , textsize = 5 ) +
  geom_signif(comparisons=list(c("inc.d9", "n9")), annotations= "p < 0.001", y_position = -15, tip_length = -0.05, vjust = 2, color= "black" , textsize = 5 )

#CORT
rechelleplot(cort.rechelle, "Corticosterone (ng/mL)")

@@

corrCORT <- hormoneswide %>%
  pivot_longer(cols = c(prolactin, estradiol, testosterone, progesterone), names_to = "hormones", values_to = "conc", values_drop_na = TRUE) %>%
  filter(study == "characterization") %>%
  ggplot(aes(x = corticosterone, y = conc)) +
  geom_point(aes(color = treatment)) +
  geom_smooth( color = "darkgrey", method = "lm") +
  facet_wrap(~hormones, nrow = 1, scales = "free_y")  +
  theme_B3() +
  theme(legend.position = "bottom",
        legend.title = element_blank()) +
  scale_color_manual(values = colorscharmaip2) +
  guides(color = guide_legend(nrow = 1)) +
  labs(y = "conc. (ng/mL)", x = "corticosterone (ng/mL)")
corrCORT
```

@@

>>>>

knitr::opts_chunk$set(fig.path = '../figures/hormones/',message=F, warning=FALSE, cache = T)
```

@@

  hormones %>% 
    filter(study == "characterization",
           hormone %in% c(myhormone))  %>% 
    droplevels() %>% 
  ggplot(aes(x = as.numeric(treatment), y = plasma_conc)) +
        geom_smooth(aes(colour = sex)) +
    geom_boxplot(aes(fill = treatment)) +
    mytheme() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
          legend.position = "none") +
    scale_fill_manual(values = colorscharmaip) +
    scale_color_manual(values = sexcolors) +
    labs(y = myylab, x = NULL) +
    guides(fill = guide_legend(order=1),
         color = guide_legend(order=2)) +
    scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9),
                       labels = charlevels)
  }

@@

@@

corrCORT <- hormoneswide %>%
  pivot_longer(cols = c(prolactin, estradiol, testosterone, progesterone), names_to = "hormones", values_to = "conc", values_drop_na = TRUE) %>%
  filter(study == "characterization") %>%
  ggplot(aes(x = corticosterone, y = conc)) +
  geom_point(aes(color = treatment)) +
  geom_smooth( color = "darkgrey", method = "lm") +
  facet_wrap(~hormones, nrow = 1, scales = "free_y")  +
  theme_B3() +
  theme(legend.position = "bottom",
        legend.title = element_blank()) +
  scale_color_manual(values = colorscharmaip2) +
  guides(color = guide_legend(nrow = 1)) +
  labs(y = "conc. (ng/mL)", x = "corticosterone (ng/mL)")
corrCORT

hormoneswide %>%
  pivot_longer(cols = c(prolactin, estradiol, testosterone, progesterone), names_to = "hormones", values_to = "conc", values_drop_na = TRUE) %>%
  filter(study == "characterization") %>%
  ggplot(aes(x = corticosterone, y = conc)) +
  geom_point(aes(color = treatment)) +
  geom_smooth(color = "darkgrey", method = "lm") +
  facet_grid(hormones ~ treatment, scales = "free_y")  +
  theme_B3() +
  theme(legend.position = "bottom",
        legend.title = element_blank()) +
  scale_color_manual(values = colorscharmaip2) +
  guides(color = guide_legend(nrow = 1)) +
  labs(y = "conc. (ng/mL)", x = "corticosterone (ng/mL)")

hormoneswide %>%
  pivot_longer(cols = c(corticosterone, estradiol, testosterone, progesterone), names_to = "hormones", values_to = "conc", values_drop_na = TRUE) %>%
  filter(study == "characterization") %>%
  ggplot(aes(x = prolactin, y = conc)) +
  geom_point(aes(color = treatment)) +
  geom_smooth(color = "darkgrey", method = "lm") +
  facet_grid(hormones ~ treatment, scales = "free_y")  +
  theme_B3() +
  theme(legend.position = "bottom",
        legend.title = element_blank()) +
  scale_color_manual(values = colorscharmaip2) +
  guides(color = guide_legend(nrow = 1)) +
  labs(y = "conc. (ng/mL)", x = "prolactin (ng/mL)")

ggplot(hormoneswide, aes(x = prolactin, y = corticosterone, color = treatment)) +
  geom_point() + geom_smooth(method = "lm") +
  facet_wrap(~treatment) +
  scale_color_manual(values =  colorscharmaip) + theme(legend.position = "none")

ggplot(hormoneswide, aes(x = prolactin, y = testosterone, color = treatment)) +
  geom_point() + geom_smooth(method = "lm") +
  facet_wrap(~treatment) +
  scale_color_manual(values =  colorscharmaip) + theme(legend.position = "none")

@@

## pca of hormones

```{r PCA}

library("factoextra")
library("FactoMineR")

# char only
hormoneswidechar <- hormoneswide %>% filter(treatment %in% charlevels)
df <- hormoneswidechar %>% select(prolactin:progesterone)
res.pca <- PCA(df)

a <- fviz_pca_var(res.pca, col.var = "black", title = "Charactherization only") 
# Contributions of variables to PC1

b <- fviz_contrib(res.pca, choice = "var", axes = 1, top = 10) + labs(y = "Dim 1 contibutions (%)",
                                                                      title = element_blank())
# Contributions of variables to PC2
c <- fviz_contrib(res.pca, choice = "var", axes = 2, top = 10) +
  labs(y = "Dim 2 contibutions (%)") + theme(title = element_blank())

plot_grid(a,b,c, nrow = 1, rel_widths = c(1,0.425,0.425), align = "hv")

# char and manip
df <- hormoneswide %>% select(prolactin:progesterone)
res.pca <- PCA(df)

a <- fviz_pca_var(res.pca, col.var = "black", title = "Charactherization & manipluation") 
# Contributions of variables to PC1

b <- fviz_contrib(res.pca, choice = "var", axes = 1, top = 10) + labs(y = "Dim 1 contibutions (%)",
                                                                      title = element_blank())
# Contributions of variables to PC2
c <- fviz_contrib(res.pca, choice = "var", axes = 2, top = 10) +
  labs(y = "Dim 2 contibutions (%)") + theme(title = element_blank())

plot_grid(a,b,c, nrow = 1, rel_widths = c(1,0.425,0.425), align = "hv")
